Coupling finite and boundary element methods to solve the Poisson-Boltzmann equation for electrostatics in molecular solvation

### 2022-04-19 methods.tex
Work on paper
Methods: adding matrix formulations and changes of notation for consistence.
Results: update of results for more refine volume mesh

The implicit solvent model can be described mathematically as a coupled system of partial differential equations, where the Poisson-Boltzmann governs in the solvent region ($\Omega_1$ in Figure XX), and the Poisson equation in the solute region ($\Omega_2$ in Figure XX). These regions are interfaced by the molecular surface ($\Gamma$), where the potential ($\phi$) and electric displacement ($\epsilon\partial\phi/\partial\mathbf{n}$) are continuous. 
%
\begin{align}\label{eq:pbe}
\nabla^2\phi_1(\mathbf{x}) &= \frac{1}{\epsilon_1}\sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k) \quad  \mathbf{x} \in \Omega_1\nonumber\\
\left(\nabla^2 - \kappa^2\right)\phi_2(\mathbf{x})  &= 0 \quad\mathbf{x}\in\Omega_2\nonumber\\
\phi_1(\mathbf{x})  = \phi_2 (\mathbf{x}), &\quad \epsilon_1\frac{\partial\phi_1}{\partial\mathbf{n}}(\mathbf{x})  = \epsilon_2\frac{\partial\phi_2}{\partial\mathbf{n}}(\mathbf{x})  \quad \mathbf{x}\in \Gamma. 
\end{align}
%
Here, $\epsilon_1$ and $\epsilon_2$ are the dielectric constants in the solute and solvent, respectively, $\kappa$ is the inverse of the Debye length, related with the salt concentration, and $q_k$ are the values of the partial charges, located at $\mathbf{x}_k$.

The electrostatic potential in $\Omega_1$ can be further decomposed into singular and regular components as $\phi_1 = \phi_c + \phi_r$, where $\phi_c$ is the solution to
%
\begin{align}\label{eq:phic}
\nabla^2\phi_c(\mathbf{x}) &= \frac{1}{\epsilon}\sum_{k=1}^{N_q}q_k\delta(\mathbf{x},\mathbf{x}_k) \quad \mathbf{x}\in\Omega_1\cup\Omega_2\nonumber\\
\phi_c(\mathbf{x})&=0 \quad \text{ as } |\mathbf{x}|\to\infty
\end{align}

@@

A common quantity of interest in implicit solvent models is the solvation free energy, which the change in Gibbs free energy as the molecule moves from vacuum into the solvent. Considering the charge distribution $\rho$ consists of point charges, this can be calculated as
%
\begin{equation}\label{eq:dG} 
\Delta G_{solv} = \frac{1}{2}\int_{\Omega_1} \rho(\mathbf{x})\phi_{r}(\mathbf{x}) = \frac{1}{2}\sum_{k=1}^{N_q} q_k\phi_r(\mathbf{x_k})
\end{equation}

@@

\subsection*{\sffamily \large BEM-BEM coupling}

@@

There are several possible boundary integral formulations of Equation \eqref{eq:pbe}.~\cite{search2022towards} The simplest form was presented by Yoon ºand Lenhoff in 1991,~\cite{YoonLenhoff1990} known as the \emph{direct} formulation, however, the resulting system is usually ill-conditioned. Here, we use the better conditioned formulation presented by Lu and co-workers.~\cite{LuETal2006,LuETal2009,debuhr2016dashmm}
Applying Green's second identity to Equation \eqref{eq:pbe} we get 
%
\begin{align} \label{eq:volume_potential}
\phi_{1}(\mathbf{r}_{\Omega_1})+ K_{L}^{\Omega_1}(\phi_{1,\Gamma}) -  V_{L}^{\Omega_1} \left(\frac{\partial}{\partial \mathbf{n}}  \phi_{1,\Gamma}  \right) & = \frac{1}{\epsilon_1} \sum_{k=0}^{N_q}  \frac{q_k}{4\pi|\mathbf{r}_{\Omega_1} - \mathbf{r}_k|}  \quad \text{on $\Omega_1$,} \nonumber \\
\phi_{2}(\mathbf{r}_{\Omega_2}) - K_{Y}^{\Omega_2}(\phi_{2,\Gamma}) + V_{Y}^{\Omega_2} \left( \frac{\partial}{\partial \mathbf{n}} \phi_{2,\Gamma} \right) & = 0 \quad \text{on $\Omega_2$,}
\end{align}
%
where $\mathbf{r}_\Omega$ is evaluated anywhere in the correspoding domain, except $\Gamma$, and $\phi_{1,\Gamma} = \phi_1(\mathbf{r}_\Gamma)$ and $\phi_{2,\Gamma} = \phi_2(\mathbf{r}_\Gamma)$ are the potentials on $\Gamma$ as we approach from $\Omega_1$ and $\Omega_2$, respectively. $K$ and $V$ are the double- and single-layer potentials for the Laplace (subscript $L$) and Yukawa (subscript $Y$, also known as modified Helmholtz) kernels
%
\begin{align}\label{eq:single_double}
V^\Omega_{L,Y}(\varphi) = \oint_\Gamma g_{L,Y}(\mathbf{r}_\Omega,\mathbf{r}')\varphi(\mathbf{r}')d\mathbf{r}'\nonumber\\
K^\Omega_{L,Y}(\varphi) = \oint_\Gamma \frac{\partial g_{L,Y}}{\partial\mathbf{n}'}(\mathbf{r}_\Omega,\mathbf{r}')\varphi(\mathbf{r}')d\mathbf{r}',\nonumber\\
\end{align}
%
where $\varphi(\mathbf{r})$ can be any distribution over $\Gamma$, and 
%
\begin{align}\label{eq:green_func}
g_L(\mathbf{r},\mathbf{r}')=\frac{1}{4\pi|\mathbf{r}-\mathbf{r}'|} \nonumber \\
g_Y(\mathbf{r},\mathbf{r}')=\frac{e^{-\kappa|\mathbf{r}-\mathbf{r}'|}}{4\pi|\mathbf{r}-\mathbf{r}'|}
\end{align}
%
are the free-space Green's function of the Laplace and linearized Poisson-Boltzmann equations, respectively.

If we take the limit of Eq. \eqref{eq:volume_potential} as $\mathbf{r}_{\Omega}$ approaches $\Gamma$, we get
%
\begin{align} \label{eq:direct}
\frac{\phi_{1,\Gamma}}{2}+ K_{L}^{\Gamma}(\phi_{1,\Gamma}) -  V_{L}^{\Gamma} \left(\frac{\partial}{\partial \mathbf{n}}  \phi_{1,\Gamma}  \right) & = \frac{1}{\epsilon_1} \sum_{k=0}^{N_q}  \frac{q_k}{4\pi|\mathbf{r}_{\Gamma} - \mathbf{r}_k|} \nonumber \\
\frac{\phi_{2,\Gamma}}{2} - K_{Y}^{\Gamma}(\phi_{2,\Gamma}) + V_{Y}^{\Gamma} \left( \frac{\partial}{\partial \mathbf{n}} \phi_{2,\Gamma} \right) & = 0
\end{align}
%
The {\it direct} formulation\cite{YoonLenhoff1990} couples the two expressions of Eq. \eqref{eq:direct} on $\Gamma$, considering the interface conditions. Even though this formulation is poorly conditioned with respect to the mesh size, it can still model large systems when it is preconditioned appropriately.\cite{AltmanBardhanWhiteTidor09,MartinezETal2019,wang2021high} 
The formulation presented by Lu\cite{LuETal2006} takes the normal derivative of \eqref{eq:direct} to couple both $\phi$ and $\partial\phi/\partial\mathbf{n}$ as
%
\begin{align}\label{eq:lu}
    & \frac{\phi_{2,\Gamma}}{2}\left(\frac{\epsilon_1}{\epsilon_2}+1\right) - \left(K_Y^\Gamma - \frac{\epsilon_1}{\epsilon_2}K_L^\Gamma\right)(\phi_{2,\Gamma}) + \left(V_Y^\Gamma - V_L^\Gamma\right)\left( \frac{\partial}{\partial \mathbf{n}} \phi_{2,\Gamma} \right) = \sum_{k=0}^{N_q}  \frac{q_k}{4\pi\epsilon_2|\mathbf{r}_{\Gamma} - \mathbf{r}_k|}
     \nonumber \\
    &\frac{\epsilon_1}{\epsilon_2}\left(W_Y^\Gamma - W_L^\Gamma\right)(\phi_{2,\Gamma}) +  \frac{1}{2}\frac{\phi_{2,\Gamma}}{\partial\mathbf{n}}\left(1+\frac{\epsilon_1}{\epsilon_2}\right) + \left(\frac{\epsilon_1}{\epsilon_2}K_Y^{\prime\Gamma} - K_L^{\prime\Gamma}\right)\left( \frac{\partial}{\partial \mathbf{n}} \phi_{2,\Gamma} \right) = \sum_{k=0}^{N_q}  \frac{\partial}{\partial\mathbf{n}_\mathbf{r}}\left(\frac{q_k}{4\pi\epsilon_2|\mathbf{r}_{\Gamma} - \mathbf{r}_k|}\right)
\end{align}
%
The two new integral operators that appear are defined as 
%
\begin{align}\label{eq:adj_hyp}
K^{\prime\Gamma}_{L,Y}(\varphi) = \oint_\Gamma \frac{g_{L,Y}}{\partial\mathbf{n}}(\mathbf{r}_\Gamma,\mathbf{r}')\varphi(\mathbf{r}')d\mathbf{r}'\nonumber\\
W^\Gamma_{L,Y}(\varphi) = - \oint_\Gamma \frac{\partial^2 g_{L,Y}}{\partial\mathbf{n}'\partial\mathbf{n}}(\mathbf{r}_\Gamma,\mathbf{r}')\varphi(\mathbf{r}')d\mathbf{r}'
\end{align}
%
and are called the {\it  adjoint double layer} and {\it hypersingular} operators, respectively. Note that the potential in Eq. \eqref{eq:lu} is evaluated on the exterior.

Having computed the electrostatic potential with Eq. \eqref{eq:lu}, we obtain the reaction potential ($\phi_r$) in Eq. \eqref{eq:dG} by subtracting out the Coulombic component out of $\phi_1$ in Eq. \eqref{eq:volume_potential}. This gives
%
\begin{align} \label{eq:phi_reac}
\phi_{r} = - K_{L}^{\Omega_1}(\phi_{1,\Gamma}) + V_{L}^{\Omega_1} \left(\frac{\partial}{\partial \mathbf{n}}  \phi_{1,\Gamma}  \right),
\end{align}
%
which is valid as long as $\epsilon_1$ is a constant.

\subsection*{\sffamily \large FEM-BEM coupling}

\begin{itemize}
    \item Description with constant and variable permittivity

@@

The coupling of finite element (FEM) and boundary element (BEM) methods is the most widely used approach for solving multi-physical problems on an unbounded domain. It allows to take advantage of both methods. On the one hand, the BEM approximates only boundary conditions, hence it is commonly used for problems involving infinite or semi-infinite domains. On the other hand, the FEM is known for its robustness and universal applicability even for problem of inhomogeneous or non-linear nature.
And that is why, we present two formualtions of FEM-BEM coupling for Poisson-Boltzmann equation.

    \item Standard FEM-BEM coupling approach

@@

    We start with the variational formulation of the internal problem. Applying integration by parts for first equation of~\eqref{eq:pbe_vp} for every $v \in H_0^1(\Omega_1)$ we have
\begin{equation}
\label{eq:fem}
 \int_{\Omega_1} \epsilon_1 \nabla \phi_1 : \nabla v ~d\mathbf{x}  - \oint_\Gamma \epsilon_1\partial_n \phi_1 v ds =   \int_{\Omega_1}  \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k)  v ~d\mathbf{x}.
\end{equation}
For external problem we use BEM and the same direct formulation as in~\eqref{eq:direct}, hence we obtain
\begin{align*}
\frac{\phi_{1,\Gamma}}{2} - K_{Y}^{\Gamma}(\phi_{1,\Gamma}) + \frac{\epsilon_1}{\epsilon_2}V_{Y}^{\Gamma} \left( \frac{\partial}{\partial \mathbf{n}} \phi_{1,\Gamma} \right) & = 0.
\end{align*}

Using the above definition and assuming that $\lambda_{1,\Gamma}  = \frac{\partial}{\partial \mathbf{n}} \phi_{1,\Gamma} $ the coupling problem can be written as follows
\begin{center}
  \textit{Find $ \phi_1 \in H^1(\Omega_1)$ and $\lambda_{1,\Gamma} \in H^{-\frac{1}{2}}(\Gamma)$ such that for all $v \in H^1(\Omega_1)$ and $\zeta \in H^{-\frac{1}{2}}(\Gamma)$}
\begin{equation} 
\label{eq:standard_fem_bem}
 \left\{
 \begin{array}{rcl}
 \int_{\Omega_1} \epsilon_1 \nabla \phi_1 : \nabla v ~d\mathbf{x}  - \oint_\Gamma \epsilon_1\partial_n \phi_1 v ds &=&   \int_{\Omega_1}  \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k)  v ~d\mathbf{x} \\[3mm] 
  \oint_\Gamma \left(\tfrac{1}{2} - K_{Y}^{\Gamma}\right) \phi_{1,\Gamma} \zeta ds + \frac{\epsilon_1}{\epsilon_2} \oint_\Gamma V_{Y}^{\Gamma} \lambda_{1,\Gamma} \zeta ds &=&0.
  \end{array}
  \right.
\end{equation}
\end{center}

%\begin{align}\label{eq:standard_fem_bem}
% \int_{\Omega_1} \epsilon_1 \nabla \phi_1 : \nabla v ~d\mathbf{x}  - \oint_\Gamma \epsilon_1\partial_n \phi_1 v ds &=   \int_{\Omega_1}  \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k)  v ~d\mathbf{x} \\
% \oint_\Gamma \frac{\phi_{1,\Gamma}}{2} - K_{Y}^{\Gamma}(\phi_{1,\Gamma}) + \frac{\epsilon_1}{\epsilon_2}V_{Y}^{\Gamma} \left( \lambda_{1,\Gamma} \right) & = 0
%\end{align}

    \item Hybrid FEM-BEM coupling

>>>>

The implicit solvent model can be described mathematically as a coupled system of partial differential equations, where the Poisson-Boltzmann governs in the solvent region ($\Omega_1$ in Figure XX), and the Poisson equation in the solute region ($\Omega_2$ in Figure XX). These regions are interfaced by the molecular surface ($\Gamma$), where the potential ($\phi$) and electric displacement ($\epsilon\partial\phi/\partial\mathbf{n}$) are continuous. 
%
\begin{align}\label{eq:pbe}
\nabla^2\phi_1(\mathbf{x}) &= \tfrac{1}{\epsilon_1}\sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k) \quad  \mathbf{x} \in \Omega_1\nonumber\\
\left(\nabla^2 - \kappa^2\right)\phi_2(\mathbf{x})  &= 0 \quad\mathbf{x}\in\Omega_2\nonumber\\
\phi_1(\mathbf{x})  = \phi_2 (\mathbf{x}), &\quad \epsilon_1\frac{\partial\phi_1}{\partial\mathbf{n}}(\mathbf{x})  = \epsilon_2\frac{\partial\phi_2}{\partial\mathbf{n}}(\mathbf{x})  \quad \mathbf{x}\in \Gamma. 
\end{align}
%
Here, $\epsilon_1$ and $\epsilon_2$ are the dielectric constants in the solute and solvent, respectively, $\kappa$ is the inverse of the Debye length, related with the salt concentration, and $q_k$ are the values of the partial charges, located at $\mathbf{x}_k$.

The electrostatic potential in $\Omega_1$ can be further decomposed into singular and regular components as $\phi_1 = \phi_c + \phi_r$, where $\phi_c$ is the solution to
%
\begin{align}\label{eq:phic}
\nabla^2\phi_c(\mathbf{x}) &= \tfrac{1}{\epsilon}\sum_{k=1}^{N_q}q_k\delta(\mathbf{x},\mathbf{x}_k) \quad \mathbf{x}\in\Omega_1\cup\Omega_2\nonumber\\
\phi_c(\mathbf{x})&=0 \quad \text{ as } |\mathbf{x}|\to\infty
\end{align}

@@

A common quantity of interest in implicit solvent models is the solvation free energy, which the change in Gibbs free energy as the molecule moves from vacuum into the solvent. Considering the charge distribution $\rho$ consists of point charges, this can be calculated as
%
\begin{equation}\label{eq:dG} 
\Delta G_{solv} = \tfrac{1}{2}\int_{\Omega_1} \rho(\mathbf{x})\phi_{r}(\mathbf{x}) = \tfrac{1}{2}\sum_{k=1}^{N_q} q_k\phi_r(\mathbf{x_k})
\end{equation}

@@

\subsubsection*{\sffamily \normalsize BEM-BEM coupling}

@@

There are several possible boundary integral formulations of Equation \eqref{eq:pbe}.~\cite{search2022towards} The simplest form was presented by Yoon ºand Lenhoff in 1991,~\cite{YoonLenhoff1990} known as the \emph{direct} formulation, however, the resulting system is usually ill-conditioned. Here, we use the better conditioned formulation presented by Lu and co-workers.~\cite{LuETal2006,LuETal2009,debuhr2016dashmm}
Applying Green's second identity to Equation \eqref{eq:pbe} we get 
%
\begin{align} \label{eq:volume_potential}
\phi_{1}(\mathbf{x}_{\Omega_1})+ K_{L}^{\Omega_1}\phi_1 -  V_{L}^{\Omega_1} \lambda_2^1 & = \tfrac{1}{\epsilon_1} \sum_{k=0}^{N_q}  \frac{q_k}{4\pi|\mathbf{x}_{\Omega_1} - \mathbf{x}_k|}  \quad \text{on $\Omega_1$,} \nonumber \\
\phi_{2}(\mathbf{x}_{\Omega_2}) - K_{Y}^{\Omega_2}\phi_2 + V_{Y}^{\Omega_2} \lambda_2& = 0 \quad \text{on $\Omega_2$,}
\end{align}
%
where $\mathbf{x}_\Omega$ is evaluated anywhere in the correspoding domain, except $\Gamma$, and $\phi_1 = \phi_1(\mathbf{x}_\Gamma)$ and $\phi_2 = \phi_2(\mathbf{x}_\Gamma)$ are the potentials on $\Gamma$ as we approach from $\Omega_1$ and $\Omega_2$, respectively. For convenience, we call the normal derivative on the interface ({\it ie.} the Neumann trace) $\lambda=\left(\frac{\partial}{\partial \mathbf{n}}  \phi_{\Gamma}  \right)$. $K$ and $V$ are the double- and single-layer potentials for the Laplace (subscript $L$) and Yukawa (subscript $Y$, also known as modified Helmholtz) kernels
%
\begin{align}\label{eq:single_double}
V^D \varphi (\mathbf{x}) = \oint_\Gamma g_i(\mathbf{x},\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'\nonumber\\
K^D \varphi (\mathbf{x}) = \oint_\Gamma \frac{\partial g_i}{\partial\mathbf{n}'}(\mathbf{x},\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}',\nonumber\\
\end{align}
%
where $D \in \{\Omega_1, \Omega_2, \Gamma\}$, $i \in \{L,Y\}$, $\varphi(\mathbf{x})$ can be any distribution over $\Gamma$, and 
%
\begin{align}\label{eq:green_func}
g_L(\mathbf{x},\mathbf{x}')=\frac{1}{4\pi|\mathbf{x}-\mathbf{x}'|} \nonumber \\
g_Y(\mathbf{x},\mathbf{x}')=\frac{e^{-\kappa|\mathbf{x}-\mathbf{x}'|}}{4\pi|\mathbf{x}-\mathbf{x}'|}
\end{align}
%
are the free-space Green's function of the Laplace and linearized Poisson-Boltzmann equations, respectively.

If we take the limit of Eq. \eqref{eq:volume_potential} as $\mathbf{x}_{\Omega}$ approaches $\Gamma$, we get
%
\begin{align} \label{eq:direct}
\tfrac{\phi_1}{2}+ K_{L}^{\Gamma}\phi_1 -  V_{L}^{\Gamma} \lambda_2^1 & = \tfrac{1}{\epsilon_1} \sum_{k=0}^{N_q}  \frac{q_k}{4\pi|\mathbf{x}_{\Gamma} - \mathbf{x}_k|} \nonumber \\
\tfrac{\phi_2}{2} - K_{Y}^{\Gamma}\phi_2 + V_{Y}^{\Gamma} \lambda_2 & = 0
\end{align}
%
The {\it direct} formulation\cite{YoonLenhoff1990} couples the two expressions of Eq. \eqref{eq:direct} on $\Gamma$, considering the interface conditions. Even though this formulation is poorly conditioned with respect to the mesh size, it can still model large systems when it is preconditioned appropriately.\cite{AltmanBardhanWhiteTidor09,MartinezETal2019,wang2021high} 
The formulation presented by Lu\cite{LuETal2006} takes the normal derivative of \eqref{eq:direct} to couple both $\phi$ and $\partial\phi/\partial\mathbf{n}$ as
%
\begin{align}\label{eq:lu}
    & \tfrac{\phi_2}{2}\left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) - \left(K_Y^\Gamma - \tfrac{\epsilon_1}{\epsilon_2}K_L^\Gamma\right)\phi_2 + \left(V_Y^\Gamma - V_L^\Gamma\right)\lambda_2 = \sum_{k=0}^{N_q}  \frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|}
     \nonumber \\
    &\tfrac{\epsilon_1}{\epsilon_2}\left(W_Y^\Gamma - W_L^\Gamma\right)\phi_2 +  \tfrac{\lambda_2}{2}\left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) + \left(\tfrac{\epsilon_1}{\epsilon_2}K_Y^{\prime\Gamma} - K_L^{\prime\Gamma}\right)\lambda_2 = \sum_{k=0}^{N_q}  \frac{\partial}{\partial\mathbf{n}_\mathbf{x}}\left(\frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|}\right)
\end{align}
%
The two new integral operators that appear are defined as 
%
\begin{align}
\label{eq:adj_hyp}
K^{\prime\Gamma}_i\varphi (\mathbf{x}) = \oint_\Gamma \frac{g_i}{\partial\mathbf{n}}(\mathbf{x}_\Gamma,\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'\nonumber\\
W^\Gamma_i\varphi (\mathbf{x}) = - \oint_\Gamma \frac{\partial^2 g_i}{\partial\mathbf{n}'\partial\mathbf{n}}(\mathbf{x}_\Gamma,\mathbf{x}')\varphi(\mathbf{x}')d\mathbf{x}'
\end{align}
where $i \in \{L,Y\}$, 
and are called the {\it  adjoint double layer} and {\it hypersingular} operators, respectively. Note that the potential in Eq. \eqref{eq:lu} is evaluated on the exterior. 
In weak form, Eq. \eqref{eq:lu} becomes
%
\begin{align}\label{eq:lu_disc}
&\left< \left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) \tfrac{\phi_{2}}{2} - \left(K_Y^\Gamma - \tfrac{\epsilon_1}{\epsilon_2}K_L^\Gamma\right)\phi_{2}, v \right>_{\Gamma} + \left< \left(V_Y^\Gamma - V_L^\Gamma\right) \lambda_2, v \right>_\Gamma = \left< \sum_{k=0}^{N_q}  f_k, v \right>_{\Gamma} \nonumber \\
&\left< \tfrac{\epsilon_1}{\epsilon_2}\left(W_Y^\Gamma - W_L^\Gamma\right)\phi_{2}, \zeta \right>_\Gamma + \left<  \left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) \tfrac{\lambda_2}{2} + \left(\tfrac{\epsilon_1}{\epsilon_2}K_Y^{\prime\Gamma} - K_L^{\prime\Gamma}\right)\lambda_2, \zeta\right>_\Gamma = \left< \sum_{k=0}^{N_q}  \frac{\partial}{\partial\mathbf{n}_\mathbf{x}} f_k, \zeta \right>_{\Gamma}
\end{align}
%
where $v, \zeta$ are test functions, $\left<\varphi,v\right>_\Gamma = \int_\Gamma \varphi(\mathbf{x})v(\mathbf{x})d\mathbf{x}$ %and $\left(\varphi,v\right)_\Omega = \int_\Omega \varphi(\mathbf{x})v(\mathbf{x})d\mathbf{x}$ are 
is the inner products on the surface% and domain, respectively
and $f_k := \frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|}$.

We introduce below the matrix formulation. Let $\vec{\phi}_2 := [\phi_2^1, \dots, \phi_2^k]^T$ be the vector of canonical basis
functions of $W_{h}^{k}$ and  $\vec{\lambda}_2 := [\lambda_2^1, \dots, \lambda_2^l]^T$ - the vector of canonical basis
functions of $\Lambda_{h}^{l}$. We define the following matrices associated with the corresponding bilinear forms
\begin{align*}
\widetilde{K}_{\alpha \beta} = \left<\left(K_Y^\Gamma - \tfrac{\epsilon_1}{\epsilon_2}K_L^\Gamma\right) \phi^{\alpha}_2, \ \lambda^{\beta}_2 \right>_{\Gamma}
 &,&
\widetilde{V}_{\alpha \beta} = \left<\left(V_Y^\Gamma - V_L^\Gamma\right) \lambda^{\alpha}_2, \ \lambda^{\beta}_2 \right>_{\Gamma}, \\
\widetilde{W}_{\alpha \beta} = \left<\left(W_Y^\Gamma - W_L^\Gamma\right) \phi^{\alpha}_2, \ \phi^{\beta}_2 \right>_{\Gamma}
 &,&
\widetilde{K'}_{\alpha \beta} = \left<\left(\tfrac{\epsilon_1}{\epsilon_2}K_Y^{\prime\Gamma} - K_L^{\prime\Gamma}\right) \lambda^{\alpha}_2, \ \phi^{\beta}_2 \right>_{\Gamma},
\end{align*}
and vectors associated with the corresponding linear forms
\begin{align*}
\vec{f}_{\beta} = \left< \sum_{k=0}^{N_q}  \frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|},   \phi^{\beta}_1  \right>_{\Gamma} &,& \vec{\partial f}_{\beta} = \left< \sum_{k=0}^{N_q}  \frac{\partial}{\partial\mathbf{n}_\mathbf{x}}\left(\frac{q_k}{4\pi\epsilon_2|\mathbf{x}_{\Gamma} - \mathbf{x}_k|}\right),   \phi^{\beta}_1  \right>_{\Gamma}.
\end{align*}

Using the above definition the discrete problem~\eqref{eq:fem_bem_disctrete} can be written in the following matrix form
\begin{align*}
\begin{bmatrix}
 \tfrac12 \left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) I - \widetilde{K}  & \widetilde{V}  \\  
 \tfrac{\epsilon_1}{\epsilon_2} \widetilde{W} & \tfrac12 \left(1+\tfrac{\epsilon_1}{\epsilon_2}\right) I + \widetilde{K'}
\end{bmatrix}
\begin{bmatrix}
\vec{\phi}_2 \\
\vec{\lambda}_2 
\end{bmatrix}
= 
\begin{bmatrix}
\vec{f} \\  
\vec{\partial f}
\end{bmatrix}.
\end{align*}
%

Having computed the electrostatic potential with Eq. \eqref{eq:lu_disc}, we obtain the reaction potential ($\phi_r$) in Eq. \eqref{eq:dG} by subtracting out the Coulombic component out of $\phi_1$ in Eq. \eqref{eq:volume_potential}. This gives
%
\begin{align} \label{eq:phi_reac}
\phi_{r} = - K_{L}^{\Omega_1}(\phi_1) + V_{L}^{\Omega_1} \left(\frac{\partial}{\partial \mathbf{n}}  \phi_1  \right),
\end{align}
%
which is valid as long as $\epsilon_1$ is a constant.

\subsection*{\sffamily \normalsize FEM-BEM coupling}

{\sffamily \small Description with constant and variable permittivity}

@@

The coupling of finite element (FEM) and boundary element (BEM) methods is the most widely used approach for solving multi-physical problems on an unbounded domain. It allows to take advantage of both methods. On the one hand, the BEM approximates only boundary conditions, hence it is commonly used for problems involving infinite or semi-infinite domains. On the other hand, the FEM is known for its robustness and universal applicability even for problem of inhomogeneous or non-linear nature.
And that is why, we present two formulations of FEM-BEM coupling for Poisson-Boltzmann equation.

{\sffamily \small Standard FEM-BEM coupling approach}

@@

    We start with the variational formulation of the internal problem. Applying integration by parts for first equation of~\eqref{eq:pbe_vp} for every $v \in H_0^1(\Omega_1)$ we have
\begin{equation}
\label{eq:fem}
 \left( \epsilon_1 \nabla \phi_1, \nabla v \right)_{\Omega_1}  -  \left<  \epsilon_1\partial_n \phi_1, v \right>_\Gamma =  \left( \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k),  v\right)_{\Omega_1},
\end{equation}
where $v$ is a test function, and $\left<\varphi,v\right>_\Gamma = \int_\Gamma \varphi(\mathbf{x})v(\mathbf{x})d\mathbf{x}$ and $\left(\varphi,v\right)_{\Omega_1} = \int_{\Omega_1} \varphi(\mathbf{x})v(\mathbf{x})d\mathbf{x}$ are the inner products on the surface and domain, respectively.
For external problem we use BEM and the same direct formulation as in~\eqref{eq:direct}, hence we obtain
\begin{align*}
\tfrac{\phi_1}{2} - K_{Y}^{\Gamma}(\phi_1) + \tfrac{\epsilon_1}{\epsilon_2}V_{Y}^{\Gamma} \left( \frac{\partial}{\partial \mathbf{n}} \phi_1 \right) & = 0.
\end{align*}

Using the above definition and assuming that $\lambda_2  = \frac{\partial}{\partial \mathbf{n}} \phi_1 $ the coupling problem can be written as follows
\begin{center}
  \textit{Find $ \phi_1 \in H^1(\Omega_1)$ and $\lambda_2 \in H^{-\frac{1}{2}}(\Gamma)$ such that for all $v \in H^1(\Omega_1)$ and $\zeta \in H^{-\frac{1}{2}}(\Gamma)$}
\begin{equation} 
\label{eq:standard_fem_bem}
 \left\{
 \begin{array}{rcl}
 \left(  \epsilon_1 \nabla \phi_1, \nabla v \right)_{\Omega_1}  - \left< \epsilon_1 \lambda_2, v \right>_\Gamma &=&   \left(  \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k),  v \right)_{\Omega_1} \\[3mm] 
  \left< \left(\tfrac{1}{2} I - K_{Y}^{\Gamma}\right) \phi_1, \zeta \right>_\Gamma + \tfrac{\epsilon_1}{\epsilon_2} \left< V_{Y}^{\Gamma} \lambda_2, \zeta \right>_\Gamma &=&0.
  \end{array}
  \right.
\end{equation}
\end{center}

%\begin{align}\label{eq:standard_fem_bem}
% \int_{\Omega_1} \epsilon_1 \nabla \phi_1 : \nabla v ~d\mathbf{x}  - \oint_\Gamma \epsilon_1\partial_n \phi_1 v ds &=   \int_{\Omega_1}  \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k)  v ~d\mathbf{x} \\
% \oint_\Gamma \tfrac{\phi_1}{2} - K_{Y}^{\Gamma}(\phi_1) + \tfrac{\epsilon_1}{\epsilon_2}V_{Y}^{\Gamma} \left( \lambda_2^1 \right) & = 0
%\end{align}

We introduce below the matrix formulation. Let $\vec{\phi}_1 := [\phi_1^1, \dots, \phi_1^j]^T$ be the vector of canonical basis
functions of the finite element space $V_{h}^{j}$, %$\vec{\phi}_2 := [\phi_2^1, \dots, \phi_2^k]^T$ - the vector of canonical basis functions of $W_{h}^{k}$,  
$\vec{\lambda}_2 := [\lambda_2^1, \dots, \lambda_2^l]^T$ - the vector of canonical basis
functions of $\Lambda_{h}^{l}$. % and $\vec{\widetilde{\phi}} := [\widetilde{\phi}^1, \dots, \widetilde{\phi}^m]^T$ be the vector of canonical basis functions of  $M_{h}^{m}$
% $T:\Omega_1 \rightarrow \Gamma$ and $E:\Gamma \rightarrow \Omega_1$ trace and extendable operators respectively.
We define the following matrices  associated with the corresponding bilinear forms
\begin{align*}
A_{\alpha \beta} = \left(\nabla \phi_1, \nabla \phi_1^{\beta} \right)_{\Omega_1}  
&,&
M_{\alpha \beta} = \left< \widetilde{\phi}^{\alpha}, \ \lambda^{\beta}_2 \right>_{\Gamma}, \\
K_{\alpha \beta} = \left<K_{Y}^{\Gamma} \phi^{\alpha}_1, \ \lambda^{\beta}_2 \right>_{\Gamma}
 &,&
V_{\alpha \beta} = \left<V_{Y}^{\Gamma} \lambda^{\alpha}_2, \ \lambda^{\beta}_2 \right>_{\Gamma}, 
\end{align*}
and vector associated with the corresponding linear form
\begin{equation*}
\vec{f}_{\beta} := \left(  \sum_{k=1}^{N_q} q_k\delta(\mathbf{x},\mathbf{x}_k),  \phi_1^{\beta} \right)_{\Omega_1}.
\end{equation*}
Using the above definition the discrete problem~\eqref{eq:standard_fem_bem} can be written in the following matrix form
\begin{align*}
\begin{bmatrix}
\epsilon_1 A &  - \epsilon_1 M^T \\  
\left(\tfrac12 I + K \right) &  \tfrac{\epsilon_1}{\epsilon_2} V 
\end{bmatrix}
\begin{bmatrix}
\vec{\phi}_1 \\  
\vec{\lambda}_2
\end{bmatrix}
= 
\begin{bmatrix}
\vec{f} \\  
0
\end{bmatrix}.
\end{align*}

{\sffamily \small Hybrid FEM-BEM coupling}
